Binary companions of red giant stars with APOGEE: \\ catalog and companions that survive the common envelope

### 2018-02-20 paper/1-catalog/twoface-catalog.tex
Some mucking about in the intro, and change jitter to excess variance.

\author{Hans-Walter~Rix}
\affiliation{Max-Planck-Institut f\"ur Astronomie,
             K\"onigstuhl 17, D-69117 Heidelberg, Germany}

% \author{Jason~Cao}
% \affiliation{Center for Cosmology and Particle Physics,
%              Department of Physics,
%              New York University, 726 Broadway,
%              New York, NY 10003, USA}

\begin{abstract}\noindent % trust me
% Context
Repeat radial-velocity measurements of stars can be used to identify stellar-,
sub-stellar, and planetary-mass companions.
Even few observations are useful for detecting companions, though such data are
difficult to use for constraining physical properties because permitted orbital
solutions are highly degenerate.
% Aims
Here we perform a search for secondary companions of stars observed by the
\apogee\ survey (\DR), which has measured multi-epoch ($\geq 3$) but sparse
radial velocities for close to 100,000 stars.
% Methods
We use a custom-built Monte Carlo sampler (\thejoker) to deliver (often highly
multi-modal) posterior beliefs about companion orbital parameters.
In some cases, where the primary-star masses are known, the mass-function
degeneracy is broken and the secondary-companion masses ($m_2\,\sin i$) can be
inferred.
% Results
We provide a catalog of \ncompanions\ companions for which the companion
properties can be confidently determined, along with posterior samplings for all
\ntotal\ stars in the parent sample.
\todo{We find differences between the companions around \RC\ and \RGB\
stars that are in/consistent with the theoretical prediction that
stellar companions should not survive the common-envelope \RGB\ phase.}
\todo{We additionally find trends in XX with chemical abundances.}

We compare the companion populations of red clump (\RC) versus \RGB\ stars at
similar surface gravity and temperature, and study the metallicity dependence
of companion properties.

\end{abstract}

@@

% Even without stellar or sub-stellar companions, the vast majority of stars have
% planetary-mass companions (\citealt{todo}).
% \todo{Planetary systems are interesting because...blah}
% Differences between the population of stellar companions and that of
% planetary-mass companions likely relate directly to the different formation
% channels for these objects.

Most of what is known about stellar companions comes from studies of
main-sequence (MS) stars.
MS stars with companions have a large dynamic range of constituent and orbital
characteristics.
For example, binary stars have mass-ratios that span from $\approx 10^-3$
\todo{citation because this is made up} to 1, and have periods from days to
millions of years.
\todo{Describe distributions over these quantities}
\todo{multiplicity in general: abundance, numbers, etc.}
\todo{some other stats from Moe and Di Stefano}

Some evolved stars have special roles in well-studied exotic multiple systems
like dwarf novae or cataclysmic variables \todo{are those examples correct?},
but by-in-large, less is known about the population of companions to evolved
stars.

- Expect differences in evolved star pop. from main sequence populations:
  - mass loss
  - envelope -> engulfment?
- Current samples limited in dynamical range of chemical properties, location,
  age (local to sun)
- Brighter, but typically larger distances, direct imaging not feasible
- Eclipsing companions good, but low mass companions stars -> hard to find
  (giants are big)
- Need large sample of stars with time-domain spectroscopy

% For planetary companions of main sequence stars, \todo{}.
% - Some indication of trends (e.g., with metallicity) -> star and
%   planet formation processes

Time-domain radial-velocity measurements of stars contain information about
massive companions: even with two successive observations of a single star, a
difference in the measured radial velocities implies the existence of at least
one companion.
However, with few or imprecise radial-velocity measurements, the orbital
properties of the companion(s) are very poorly constrained (e.g.,
\citealt{Price-Whelan:2017}).
The vast majority of spectroscopic targets with repeat observations in the
largest (by number of objects) stellar spectroscopic surveys are often observed
just a few times with sparse, non-uniform phase coverage.
Most prior searches for companions using survey RV data have therefore
restricted their searches to only sources with many, high-precision epochs, so
that the orbital solution can be unambiguously determined (e.g.,
\citealt{Troup:2016}), or have used simple statistics computed from the data to
study multiplicity (e.g., $\textrm{RV}_\textrm{max} - \textrm{RV}_\textrm{min}$;
\citealt{Badenes:2017}).

@@

We have solved this problem previously, though with limitations (to be discussed
more below), with \thejoker\ (\citealt{Price-Whelan:2017}).
\thejoker\ is a Monte Carlo rejection sampler that is computationally expensive
but probabilistically righteous:
It delivers independent posterior pdf samples for single-companion binary
orbital parameters, given any number of radial-velocity measurements.
Here we use \thejoker\ to generate posterior pdf samples for stars observed by
the \apogee\ survey (see \sectionname~\ref{sec:data}; \citealt{Majewski:2015}).

@@

All data used in this work comes from the publicly-available data release 14
(\DR) of the \apogee\ survey (\citealt{apogeedr14}), a component of the Sloan
Digital Sky Survey IV (\sdssiv; \citealt{sdss,Blanton:2017}).
\apogee\ is designed to map stars across much of the Milky Way by obtaining
high-resolution ($R \sim 22500$) infrared ($H$-band) spectroscopy of primarily
\RGB\ stars.
Targets are selected with simple color and brightness cuts, but the survey uses
fiber-plugged plates with a maximum of 300 fibers per each $\approx
1.5~\textrm{deg}^2$ field of view, leading to ``pencil-beam''-like sampling of
the stellar distribution.
Data taken as part of the \apogee\ survey is reduced with the \apogee\ Stellar
Parameters and Chemical Abundances Pipeline (\acronym{ASPCAP};
\citealt{Garcia-Perez:2016}), which measures stellar parameters, chemical
abundances, and radial velocities for each target.
In order to meet signal-to-noise ratio requirements, most \apogee\ stars are
observed multiple times in a series of ``visits,'' typically with at least one
visit separated by a month or more in order to help identify binary stars.
\acronym{ASPCAP} iteratively solves for the relative velocities of each visit
spectrum while determining the optimal co-addition of the visit spectra.

We use the primary data products from \apogee\ \DR\ (i.e. the \texttt{allStar}
and \texttt{allVisit} files) which contain 258,475 unique source IDs
(\texttt{APOGEE\_ID}) and 1,054,381 unique visits.
We select all stars with $\geq 3$ visits that each pass a set of quality cuts,
described below.
For each visit, we require that the visit velocity uncertainty is $< 100~\kms$
(\texttt{VRELERR}) and the following bits are not set in the \texttt{STARFLAGS}
bitmask: \texttt{PERSIST\_HIGH}, \texttt{PERSIST\_JUMP\_POS},
\texttt{PERSIST\_JUMP\_NEG}, \texttt{VERY\_BRIGHT\_NEIGHBOR}, \texttt{LOW\_SNR}.
For each star, we require that $0 < \log g < 3.5$ and the following bits are not
set in the \texttt{ASPCAPFLAGS} bitmask: \texttt{STAR\_BAD}.
After these cuts, and the requirement of $\geq 3$ visits for a given star, we
are left with \nvisits\ visits for \nstars\ unique sources.
\figurename~\ref{fig:nvisits} shows the number of stars in several bins of
number of visits that pass the above quality cuts: \todo{XX}\% of the stars in
\DR\ have $< 8$ visits.

% Notebook: "Numbers after cuts"
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.7\textwidth]{nvisits.pdf}
\end{center}
\caption{%
\todo{Update with new apogee db file}
Number of \apogee\ stars in logarithmic bins of number of visits that pass the
quality cuts described in \sectionname~\ref{sec:data}.
In total, this work uses \nvisits\ visits and \nstars\ unique sources.
\label{fig:nvisits}
}
\end{figure}

@@

We perform our individual system fits---that is, make our posterior
samplings---under the following assumptions:
\begin{description}
\item[no multiplets] All radial-velocity variations of the primary star are
  induced by a single companion.
  This is motivated by the idea that triple star systems are usually
  hierarchical so that the period of the inner binary is typically much shorter
  than the orbital period of the outer body (\todo{CITATION NEEDED}).
  At present, we ignore the possibility of higher-order multiple systems.
  % \todo{revisit this - do we want to also sample over a long-term trend?}
\item[Kepler] Related to the first point, all velocity variations of the primary
  are gravitational, and we therefore ignore the possibility of coherent intrinsic variation from, e.g., stellar oscillations.
\item[SB1] All spectra are single-lined; that is, we assume that the secondary
  is significantly fainter and is thus undetected in the spectra.
  This assumption is motivated by the fact that we expect \RGB\ stars to be
  substantially more luminous than their typical companion.
  However, there are known double-lined binary stars in the \apogee\ catalog
  \citealt{El-Badry:2017}, and an expected but unknown fraction of \RGB--\RGB\
  binaries.
\item[simple noise model] Measurements are unbiased and noise estimates are
  correct up to an unknown extra variance.
  All noise contributions result in Gaussian uncertainties on the individual
  radial-velocity measurements.
  We often refer to the extra variance as a ``jitter'' parameter because for
  upper \RGB\ stars the inferred extra variance will be a combination of extra
  uncertainty and true astrophysical surface jitter (\todo{CITATION NEEDED}).
\end{description}

@@

\subsection{Model selection}

In addition to generating posterior samples for each source, we also compute the
fully marginalized likelihood (FML) under two other models: (1) a model in which
the radial velocity of the source is constant, and (2) a model in which the
radial-velocity variations are purely linear.
We later use these FML values to define cuts on our full catalog of posterior
samples to produce a catalog of companions.

\todo{equations?}

Examples of outputs and etc.

@@

For all \nstars\ \apogee\ stars with $\geq 3$ good visits (see
\sectionname~\ref{sec:data}), we use \thejoker\ to generate posterior samplings
for each star under the assumptions listed above (see
\sectionname~\ref{sec:fitting}).
We start by generating a library of \nprior\ prior samples generated under a
prior similar to that defined in \citet{Price-Whelan:2017}: uniform or isotropic
in angle parameters, uniform in log-period over the domain $[1,
32768]~\textrm{day}$, and a beta distribution over eccentricity with fixed
parameters (\citealt{Kipping:2013}).
For the prior over the extra variance or jitter parameter, $s$, we initially use
a Gaussian over the transformed parameter $y = \ln s^2$ with the mean and
standard deviation of the initial distribution set to $(\mu_{y,0}, \sigma_{y,0})
= (10.6, 3)$ in units of $\textrm{m}~\textrm{s}^{-1}$.

@@

\subsection{Inferring the extra variance distribution}
\label{sec:inferjitter}

As an initial use of the per-source posterior samplings, we use a hierarchical
Bayesian model to infer the parameters of the (assumed Gaussian) distribution
over the log-jitter parameter: $(\mu_y, \sigma_y)$.
This inference serves as a test-case for future work, where we intend to use the
independent posterior samplings to construct a hierarchical inference over
companion property distributions.
This is also a test of the visit velocity uncertainties reported in the \apogee\
data products: If the catalog uncertainties are significantly underestimated, we
expect the inferred log-jitter distribution parameters to tend towards larger
values.

@@

For each $n$ of $N$ RGB stars in APOGEE, we obtain $K$ posterior samples over
primary orbital parameters $\bs{\theta} = (P, e, \omega, M_0, K, v_0)$ and the
jitter parameter, $y = \ln s^2$, using \thejoker; For brevity in expressions
below, we will use the vector
\begin{equation}
    \bs{w} = (\bs{\theta}, y)
\end{equation}
to represent the full set of parameters.
To obtain this sampling, we use an interim (Gaussian) prior on the jitter
parameter parametrized by a mean and standard deviation, i.e. $\alpha_0 =
(\mu_{y,0}, \sigma_{y,0})$ as described above.
For a given source, the posterior samples in the above parameters are drawn from
the distribution
\begin{equation}
    \bs{w}_k \sim p(\bs{w}_k \given D, \bs{\alpha}_0)
\end{equation}
where $D$ represents the data for a given object.

We want to compute the likelihood of all data from all $N$ stars, $\{D_n\}$,
given a new set of hyperparameters $\bs{\alpha}$
\begin{equation}
    p(\{D_n\} \given \bs{\alpha}) = \prod_n^N p(D_n \given \bs{\alpha})
\end{equation}
where in the above, we have assumed that this likelihood is separable ( the data
for each source are independent).
The per-source marginal likelihood in the above expression is given by
\begin{align}
    p(D_n \given \bs{\alpha}) &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \,
      p(\bs{w}_n \given \bs{\alpha})\\
    &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \, p(\bs{w}_n \given \bs{\alpha}) \,
      \frac{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}\\
    &= p(D_n \given \bs{\alpha}_0) \, \int \dd \bs{w}_n \,
      \frac{p(\bs{w}_n \given \bs{\alpha})}{p(\bs{w}_n \given \bs{\alpha}_0)} \,
      p(\bs{w}_n \given D_n, \bs{\alpha}_0) \label{eq:marglike} \quad .
\end{align}
Using the Monte Carlo integration approximation, \eqname~\ref{eq:marglike} can
be simplified to a sum over prior value ratios of the $K$ posterior samples in
the log-jitter parameter for each $n$ star
\begin{equation}
    \approx \frac{\mathcal{Z}_n}{K} \,
      \sum_k^{K} \frac{p(y_{nk} \given \bs{\alpha})}{p(y_{nk} \given \bs{\alpha}_0)}
\end{equation}
where we have canceled the other priors (over $\bs{\theta}$), and all
normalization constants appear in the constant scale factor $\mathcal{Z}_n$.

@@

We use 1,825 stars with $>10$ visits and $\log g < 2$ (to avoid upper RGB stars
that have large intrinsic jitter) and maximize the above likelihood to determine
better hyperparameters for the log-jitter parameter distribution.
\figurename~\ref{fig:infer-jitter} shows the distribution corresponding to the
maximum-likelihood hyperparameters, $\alpha^* = (\mu_y^*, \sigma_y^*) = (9.50,
1.64)$.
\todo{discussion about this}
\todo{some discussion of how this relates to empirical bayes?}

>>>>

\author[0000-0003-4996-9069]{Hans-Walter~Rix}
\affiliation{Max-Planck-Institut f\"ur Astronomie,
             K\"onigstuhl 17, D-69117 Heidelberg, Germany}

\author{APOGEE Team}

\begin{abstract}\noindent % trust me
% Context
Repeat radial-velocity measurements of stars can be used to identify stellar,
sub-stellar, and planetary-mass companions.
Even sparse or few-epoch observations are useful for detecting companions,
though such data are difficult to use for measuring physical properties of
individual systems because permitted orbital solutions can be highly degenerate.
% Aims
Using a custom-built Monte Carlo sampler (\thejoker) to deliver (often highly
multi-modal) posterior beliefs about companion orbital parameters, here we
perform a search for secondary companions of over 100,000 stars observed by the
\apogee\ survey (\DR) with $\geq 3$ epochs.
% Methods
We select stars with probable companions by \todo{how?}.
% Results
We provide (1) a catalog of \ncompanions\ companions for which the companion
properties can be confidently determined, (2) a catalog of stars that likely
have companions, but for  which more observations are needed to uniquely
determine the orbital properties, and (3) posterior samplings for all \ntotal\
stars in the parent sample.
We highlight interesting systems and show the characteristics of stars with
confidently determined companion properties.
\end{abstract}

@@

Most of what is known about stellar companions comes from studies of
nearby main-sequence (MS) stars.
MS stars with companions have a large dynamic range of constituent and orbital
characteristics.
For example, binary stars have mass-ratios that span from $\approx 0.03$
to 1 (e.g., \citealt{Kraus:2008}), and have periods from days to
millions of years (e.g., \citealt{Raghavan:2010}).
\todo{multiplicity in general: occurrence rates}

Less is known about uninteracting or detached companions to evolved stars.
Such systems are interesting because they enable studying extreme outcomes or
conditions of star formation and multiplicity.
For example, evolved stars are typically much brighter than their MS
counterparts, which enables characterizing the stellar companion population
throughout the Galactic disk over a wide range of chemical abundances.
They also provide a means to study the outcomes of stellar engulfment or
common-envelope evolution: Do short-period companions exist around evolved
stars, and do their statistics constrain stellar evolution on the red giant
branch (RGB)?
\todo{motivate time-domain spectroscopy}

Time-domain radial-velocity measurements of stars contain information about
massive companions: even with two successive observations of a single star, a
difference in the measured radial velocities implies the existence of at least
one companion.
However, with few or imprecise radial-velocity measurements, the orbital
properties of the companion(s) are very poorly constrained (e.g.,
\citealt{Price-Whelan:2017}).
The vast majority of spectroscopic targets with repeat observations in the
largest (by number of objects) stellar spectroscopic surveys are often observed
just a few times with sparse, non-uniform phase coverage.
Most prior searches for companions using survey RV data have therefore
restricted their searches to only sources with many, high-precision epochs, so
that the orbital solution can be unambiguously determined (e.g.,
\citealt{Troup:2016}), or have used simple statistics computed from the data to
study multiplicity (e.g., $\textrm{RV}_\textrm{max}$; \citealt{Badenes:2017}).

@@

We have solved this problem previously, though with limitations (to be discussed
more below), with \thejoker\ (\citealt{Price-Whelan:2017}).
\thejoker\ is a Monte Carlo rejection sampler that is computationally expensive
but probabilistically righteous:
It delivers independent posterior pdf samples for single-companion binary
orbital parameters, given any number of radial-velocity measurements.
Here we use \thejoker\ to generate posterior pdf samples for stars observed by
the \apogee\ survey (see \sectionname~\ref{sec:data}; \citealt{Majewski:2017}).

@@

All data used in this work comes from the publicly-available data release 14
(\DR) of the \apogee\ survey (\citealt{Majewski:2017,Abolfathi:2017}), a
component of the Sloan Digital Sky Survey IV (\sdssiv;
\citealt{Gunn:2006,Blanton:2017}).
\apogee\ is designed to map stars across much of the Milky Way by obtaining
high-resolution ($R \sim 22,500$) infrared ($H$-band) spectroscopy of primarily
\RGB\ stars.
Targets are selected with simple color and brightness cuts, but the survey uses
fiber-plugged plates with a maximum of 300 fibers per each $\approx
1.5~\textrm{deg}^2$ field of view, leading to ``pencil-beam''-like sampling of
the stellar distribution.
In order to meet signal-to-noise ratio requirements, most \apogee\ stars are
observed multiple times in a series of ``visits,'' typically with at least one
visit separated by a month or more in order to help identify binary stars.

Data taken as part of the \apogee\ survey is reduced with a multi-step data
reduction pipeline that ultimately solves for the stellar parameters, chemical
abundances, and radial velocities for each target
(\citealt{Nidever:2015}).
Most relevant for this work, the visit radial velocities (RVs) are determined
using an iterative scheme: the individual visit spectra are combined using
initial guesses for the relative RVs into a coadded spectrum, which is then used
to re-derive the relative visit velocities.
The stellar parameters---surface gravity, $\log g$, and effective temperature,
$T_{\textrm{eff}}$---and the chemical abundances are determined from the coadded
spectrum as a part of the \apogee\ Stellar Parameters and Chemical Abundances
Pipeline (\acronym{ASPCAP}; \citealt{ASPCAP}).

We use the primary data products from \apogee\ \DR\ (i.e. the \texttt{allStar}
and \texttt{allVisit} files) which contain 258,475 unique source IDs
(\texttt{APOGEE\_ID}) and 1,054,381 unique visits.
We select all stars with $\geq 3$ visits that each pass a set of quality cuts,
described below.
For each visit, we require that the visit velocity uncertainty is $< 100~\kms$
(\texttt{VRELERR}) and the following bits are not set in the \texttt{STARFLAGS}
bitmask: \texttt{PERSIST\_HIGH}, \texttt{PERSIST\_JUMP\_POS},
\texttt{PERSIST\_JUMP\_NEG}, \texttt{VERY\_BRIGHT\_NEIGHBOR}, \texttt{LOW\_SNR}.
For each star, we require that $0 < \log g < 4$ and the following bits are not
set in the \texttt{ASPCAPFLAGS} bitmask: \texttt{STAR\_BAD}.
After these cuts, and the requirement of $\geq 3$ visits for a given star, we
are left with \nvisits\ visits for \nstars\ unique sources.
\figurename~\ref{fig:nvisits} shows the number of stars in several bins of
number of visits that pass the above quality cuts: \todo{XX}\% of the stars in
\DR\ have $< 8$ visits.
\figurename~\ref{fig:loggteff} ...

% Notebook: "Numbers of visits and sample overview"
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.7\textwidth]{nvisits.pdf}
\end{center}
\caption{%
Number of \apogee\ stars in logarithmic bins of number of visits that pass the
quality cuts described in \sectionname~\ref{sec:data}.
In total, this work uses \nvisits\ visits and \nstars\ unique sources.
\label{fig:nvisits}
}
\end{figure}

% Notebook: "Numbers of visits and sample overview"
\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{logg_teff_feh.pdf}
\end{center}
\caption{%
\textit{Left panel}: Number of stars in bins of iron abundance,
$[\textrm{Fe}/\textrm{H}]$, that pass the quality cuts described in
\sectionname~\ref{sec:data}.
\textit{Right panel}: Distribution of stars in our sample in stellar parameters
log-surface-gravity, $\log g$, and effective temperature, $T_{\textrm{eff}}$,
with points colored by the iron abundance.
\label{fig:loggteff}
}
\end{figure}

@@

We perform our individual system fits---that is, make our posterior
samplings---under the following assumptions:
\begin{description}
\item[no multiplets] All radial-velocity variations of the primary star are
  induced by a single companion.
  This is motivated by the idea that triple star systems are usually
  hierarchical so that the period of the inner binary is typically much shorter
  than the orbital period of the outer body (\todo{CITATION NEEDED}).
  At present, we ignore the possibility of higher-order multiple systems.
  % \todo{revisit this - do we want to also sample over a long-term trend?}
\item[Kepler] Related to the first point, all velocity variations of the primary
  are gravitational, and we therefore ignore the possibility of coherent intrinsic variation from, e.g., stellar oscillations.
\item[SB1] All spectra are single-lined; that is, we assume that the secondary
  is significantly fainter and is thus undetected in the spectra.
  This assumption is motivated by the fact that we expect \RGB\ stars to be
  substantially more luminous than their typical companion.
  However, there are known main-sequence double-lined binary stars in the
  \apogee\ catalog \citealt{El-Badry:2018}, and an expected but unknown fraction
  of \RGB--\RGB\ binaries.
\item[simple noise model] Measurements are unbiased and noise estimates are
  correct up to an unknown excess variance.
  All noise contributions result in Gaussian uncertainties on the individual
  radial-velocity measurements.
  For upper \RGB\ stars the inferred excess variance will be a combination of
  extra systematic uncertainty and true astrophysical surface jitter
  (\todo{CITATION NEEDED}).
\end{description}

@@

% \subsection{Model selection}
%
% In addition to generating posterior samples for each source, we also compute the
% fully marginalized likelihood (FML) under two other models: (1) a model in which
% the radial velocity of the source is constant, and (2) a model in which the
% radial-velocity variations are purely linear.
% We later use these FML values to define cuts on our full catalog of posterior
% samples to produce a catalog of companions.
%
% \todo{equations?}
%
% Examples of outputs and etc.

@@

For all \nstars\ \apogee\ stars with $\geq 3$ good visits (see
\sectionname~\ref{sec:data}), we use \thejoker\ to generate posterior samplings
for each star under the assumptions listed above (see
\sectionname~\ref{sec:fitting}).
We start by generating a library of \nprior\ prior samples generated under a
prior similar to that defined in \citet{Price-Whelan:2017}:
\begin{itemize}
    \item uniform or isotropic in angle parameters,
    \item uniform in log-period over the domain $[1,32768]~\textrm{day}$,
    \item a beta distribution over eccentricity with fixed parameters (\citealt{Kipping:2013}).
\end{itemize}
For the prior over the excess variance parameter, $s$, we initially use
a Gaussian over the transformed parameter $y = \ln s^2$ with the mean and
standard deviation of the initial distribution set to $(\mu_{y,0}, \sigma_{y,0})
= (10.6, 3)$ in units of $\textrm{m}~\textrm{s}^{-1}$.

@@

\subsection{Inferring the excess variance distribution}
\label{sec:inferjitter}

As an initial use of the per-source posterior samplings, we use a hierarchical
Bayesian model to infer the parameters of the (assumed Gaussian) distribution
over the log-excess-variance parameter: $(\mu_y, \sigma_y)$.
This inference serves as a test-case for future work, where we intend to use the
independent posterior samplings to construct a hierarchical inference over
companion property distributions.
This is also a test of the visit velocity uncertainties reported in the \apogee\
data products: If the catalog uncertainties are significantly underestimated, we
expect the inferred log-excess-variance distribution parameters to tend towards
larger values.

@@

For each $n$ of $N$ RGB stars in APOGEE, we obtain $K$ posterior samples over
primary orbital parameters $\bs{\theta} = (P, e, \omega, M_0, K, v_0)$ and the
excess-variance parameter, $y = \ln s^2$, using \thejoker; For brevity in
expressions below, we will use the vector
\begin{equation}
    \bs{w} = (\bs{\theta}, y)
\end{equation}
to represent the full set of parameters.
To obtain this sampling, we use an interim (Gaussian) prior on the
excess-variance parameter parametrized by a mean and standard deviation, i.e.
$\alpha_0 = (\mu_{y,0}, \sigma_{y,0})$ as described above.
For a given source, the posterior samples in the above parameters are drawn from
the distribution
\begin{equation}
    \bs{w}_k \sim p(\bs{w}_k \given D, \bs{\alpha}_0)
\end{equation}
where $D$ represents the data for a given object.

We want to compute the likelihood of all data from all $N$ stars, $\{D_n\}$,
given a new set of hyperparameters $\bs{\alpha}$
\begin{equation}
    p(\{D_n\} \given \bs{\alpha}) = \prod_n^N p(D_n \given \bs{\alpha})
\end{equation}
where in the above, we have assumed that this likelihood is separable ( the data
for each source are independent).
The per-source marginal likelihood in the above expression is given by
\begin{align}
    p(D_n \given \bs{\alpha}) &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \,
      p(\bs{w}_n \given \bs{\alpha})\\
    &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \, p(\bs{w}_n \given \bs{\alpha}) \,
      \frac{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}\\
    &= p(D_n \given \bs{\alpha}_0) \, \int \dd \bs{w}_n \,
      \frac{p(\bs{w}_n \given \bs{\alpha})}{p(\bs{w}_n \given \bs{\alpha}_0)} \,
      p(\bs{w}_n \given D_n, \bs{\alpha}_0) \label{eq:marglike} \quad .
\end{align}
Using the Monte Carlo integration approximation, \eqname~\ref{eq:marglike} can
be simplified to a sum over prior value ratios of the $K$ posterior samples in
the log-excess-variance parameter for each $n$ star
\begin{equation}
    \approx \frac{\mathcal{Z}_n}{K} \,
      \sum_k^{K} \frac{p(y_{nk} \given \bs{\alpha})}{p(y_{nk} \given \bs{\alpha}_0)}
\end{equation}
where we have canceled the other priors (over $\bs{\theta}$), and all
normalization constants appear in the constant scale factor $\mathcal{Z}_n$.

@@

We use 1,825 stars with $>10$ visits and $\log g < 2$ (to avoid upper RGB stars
that have large intrinsic jitter) and maximize the above likelihood to determine
better hyperparameters for the log-excess-variance parameter distribution.
\figurename~\ref{fig:infer-jitter} shows the distribution corresponding to the
maximum-likelihood hyperparameters, $\alpha^* = (\mu_y^*, \sigma_y^*) = (9.50,
1.64)$.
\todo{discussion about this}
\todo{some discussion of how this relates to empirical bayes?}
